Online Balanced Graph Partitioning: Optimizations for Large Clusters

### 2020-05-06 paper/disc20clustering.tex
wip

 \newcommand\mahmoud[1]{\color{orange}\textbf{\\ Mahmoud: #1}\\\color{black}}
 \newcommand\stefan[1]{\color{blue}\textbf{\\ Stefan: #1}\color{black}}
 \newcommand\maciek[1]{\color{brown}\textbf{\\ Maciek: #1}\color{black}}
 %\newcommand\maciek[1]{}

@@

The algorithm for $k=3$ from Section~\ref{sec:k3} can be efficiently distributed.
There are two non-local operations:
\begin{enumerate}
	\item We broadcast the end of the phase to reset all counters.
	\item Merge of two  may require finding a third cluster that contains $3$ alone nodes (For the details, see the proof of Lemma~\ref{lem:1req})).
\end{enumerate}

If a component grows to size $4$ or fresh cluster is found, we conclude that no perfect partition exists, and we broadcast (\maciek{TODO: efficient broadcast cite}) the message to end the phase and reset the counters in all clusters.

Let's call the cluster containing $3$ alone nodes \emph{fresh}.
Now we show how to efficiently find a fresh cluster.
We organize the clusters into an arbitrary balanced binary tree (\maciek{TODO: how to do it efficiently? Does not matter much, as this is performed once at the beginning}), and broadcast the root.
Each cluster counts the number of fresh clusters in its subtree.
To find a fresh cluster, we traverse an arbitrary path of non-zero counters from the root.
Upon encountering a fresh cluster, we stop the traversal and decrease the counters on the followed path by $1$.

The operation (ref1) requires a single broadcast, and the operation (ref2) requires $O(\log(\ell))$ communication.
Each of these may occur on every request.

\maciek{Possible improvement: we can have a fixed number of rounds for finding the X cluster by maintaining a direct pointer to an X cluster (not only to subgroup). However, the fix still requires the log number of rounds. But we improved the responsiveness.}

@@

\pagebreak
\appendix

\section{Omitted results}

\begin{enumerate}
	\item The lower bound of $\Omega(\ell)$ for the scenario with at most $(1/3-1/k)$-augmentation

	\item Costly cascade example (it is easy to present $k^2$ and $k\cdot \sqrt{k}$ (with different approach), and then we can combine these to obtain $k^3$)

@@

	\item lower and upper bound $k+1$ for two clusters, perfect partition and the optimal configuration reachable within one swap
	\item 
	The bound $(k \cdot \ell - k - \ell) / 2$ can be easily improved to $(k \cdot \ell - k - 1) / 2$ by playing a simple caching game in the first cluster, instead of just giving away the $k-1$ "blocker" (the caching game would build it gradually)

>>>>

 \newcommand\mahmoud[1]{\color{orange}\textbf{\\ Mahmoud: #1}\\\color{black}}
 \newcommand\stefan[1]{\color{blue}\textbf{\\ Stefan: #1}\color{black}}
 \newcommand\maciek[1]{\color{brown}\textbf{(Maciek: #1)}\color{black}}
 %\newcommand\maciek[1]{}

@@

The algorithm for $k=3$ from Section~\ref{sec:k3} can be efficiently distributed.
There are two non-local operations:
\begin{enumerate}
	\item Broadcasting the end of the phase.\label{it:broadcast}
	\item Component merge.
	\label{it:merge}
\end{enumerate}

If a component reaches size $4$ or fresh cluster is not found, we conclude that no perfect partition exists.
Then, we end the phase and broadcast (\maciek{TODO: efficient broadcast cite}) the message to  reset the counters in all clusters.

A merge of two components may require finding a third cluster that contains $3$ alone nodes (for details see the proof of Lemma~\ref{lem:1req})).
We say that a cluster containing $3$ alone nodes is \emph{fresh}.
In the following, we show how to efficiently find a fresh cluster in a distributed manner.
We organize the clusters into an arbitrary rooted balanced binary tree (\maciek{TODO: how to do it efficiently? Does not matter much, as this is performed once at the beginning}), and we broadcast the root to each cluster.
Each cluster maintains the counter of fresh clusters in its subtree.
To find a fresh cluster, we traverse an arbitrary path of non-zero counters from the root.
Upon encountering a fresh cluster, we stop the traversal and decrease the counters on the followed path by $1$.

The operation (\ref{it:broadcast}) requires a single broadcast, and the operation (\ref{it:merge}) requires $O(\log(\ell))$ communication.
Each of these may occur, on average, every $\alpha$ external requests.

@@

\appendix

\section{Omitted results (for our future papers)}

\begin{enumerate}
	\item Costly cascade example (it is easy to present $k^2$ and $k\cdot \sqrt{k}$ (with different approach), and then we can combine these to obtain $k^3$)

@@

	\item lower and upper bound $k+1$ for two clusters, perfect partition and the optimal configuration reachable within one swap

	\item The lower bound $(k \cdot \ell - k - \ell) / 2$ can be easily improved to $(k \cdot \ell - k - 1) / 2$ by playing a simple caching game in the first cluster, instead of just giving away the $k-1$ "blocker" (the caching game would build it gradually)
